---
title: "Horse Colic Project"
author: Destinee Morrow
output: ioslides_presentation
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE)
library(dplyr)
library(ggplot2)
library(tidyverse)
```

## Background
Colic - pain in the abdomen of the horse.

Various types of horse colic: 

- impaction colic due to the type of feed and the eating process 
- gas colic that is built up over time
    - can lead to more serve colic, like abdomen dystension and colon shift
    
All of these, can potentially lead to death if the horse if not treated. Horse colic is very common and understanding the severity and needed procedures is important.


## Data: {.smaller}
```{r Horse Colic, echo=FALSE}
setwd("/Users/destinee/Documents/Programming")
horsecolic <- read.csv("horsecolic.csv")
```

```{r table, echo=FALSE}
head(horsecolic)
```

## Packed Cell Volume

```{r horse PCV, echo=FALSE}
ggplot(horsecolic %>% drop_na(packed_cell_volume), aes(packed_cell_volume)) + geom_density( aes(fill = packed_cell_volume < 40), alpha=0.5) + scale_fill_manual(values = c("blue","red"), 0.2) + xlab("Packed Cell Volume") +ylab("Number of Horses") +ggtitle("Number of Colic Horses as a function of PCV")
```


## PCV T-test
```{r ttest, echo=FALSE}
t.test(horsecolic$packed_cell_volume ~ horsecolic$surgery)
```
Fail to reject the null hypothesis. 
So... If PCV cant tell us whether or not the horse will have to have surgery.
The question is...
Should the horse ultimately have surgery or not?

## Surgery
 
```{r horse facet, echo=FALSE}
horsecolic[horsecolic$outcome=='euthanized', "outcome2"] <- "died"
horsecolic[horsecolic$outcome=='lived', "outcome2"] <- "lived"
horsecolic[horsecolic$outcome=='died', "outcome2"] <- "died" 
hoc <- horsecolic %>% group_by(surgery) %>% count(outcome)
hoc <- as.data.frame(hoc)
ggplot(hoc, aes(outcome, n)) + geom_col(aes(fill=outcome)) +ylim(0,110) + facet_grid(. ~ surgery) +xlab("Outcome/Survival of Horses") + ylab("Number of Horses") + ggtitle("Outcome of Colic Horses as a function of Receiving Surgery")
```


## Surgery 2

```{r, echo=FALSE}
hoc <- horsecolic %>% group_by(surgery) %>% count(outcome2)
hoc <- as.data.frame(hoc)
ggplot(hoc, aes(outcome2, n), na.rm=TRUE) + geom_col(aes(fill=outcome2)) +ylim(0,130) + facet_grid(. ~ surgery) +xlab("Outcome/Survival of Horses") + ylab("Number of Horses") + ggtitle("Outcome of Colic Horses as a function of Receiving Surgery")
```

So should we have a horse undergo surgery to fix the colic?
We will use a Chi-squared Test to determine that answer.

## Chi-Squared

```{r chisq, echo=FALSE}
tbl <- table(horsecolic$surgery, horsecolic$outcome)
chisq.test(tbl)
```

```{r chisqlive, echo=FALSE}
tbl2 <- table(horsecolic$surgery, horsecolic$outcome2)
chisq.test(tbl2)
```

## Conclusion
The Chi-Squared tests are based on a null hypothesis of surgery and survival being independent. 
The p values of 0.004424 and 0.005016 are significantly smaller than that of the 0.05 alpha. Thus meaning, that we reject the null hypothesis that whether a horse has surgery or not is independent of whether the horse survives from colic. 

These two variables are dependent and so, no, we should not perform surgery on a horse that is experiencing colic because it is more likely to die than survive.

