---
title: "Week 5 Assignment"
author: "Caleb Timmons"
date: "2/18/2020"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## First Dataset

Source: <https://datadryad.org/stash/dataset/doi:10.5061/dryad.00t8g>

This dataset examines the effects of several treatments on the rate of honey bee visits.

The main problem is there is some manipulation necessary to get the data fully ready for analysis.  Many integers need to be converted to factors among other steps.
```{r}
bees <- read.csv('bees.csv', header = TRUE)
bees_df <- bees[,c("Plant_Code","Bee_Count","Temp_F","Wind_mph",'Barometric','Humidity')]
summary(bees_df)
```
```{r, echo=FALSE}
boxplot(bees$Bee_Count ~ bees$Sugar,
        main = 'Beecounts for Sugar Treatment',
        xlab = 'Sugar Treatment',
        ylab = 'Bee Counts')
```

## Second Dataset

Source: <http://mlr.cs.umass.edu/ml/datasets/Abalone>

This dataset seeks to determine if there is a method of determining the age of an Abalone from physical measurments as opposed to ring counting in the shell.

There are variable for Sex, Length, Diameter, Height, Whole Weight, Shucked Weight, Viscera Weight, Shell Weight, and rings.

One potential problem is all measurements have been divided by 200.  Analysis may requrie undoing this.

A possible question is determining a formula that can be used to determine the age given measurements.

```{r}
abalone <- read.csv('abalone.data.txt', header = TRUE)
summary(abalone)
```
```{r, echo=FALSE}
plot(abalone$Len~abalone$Ring,
     main = 'Abalone Length vs Ring Number',
     xlab = 'Ring Number',
     ylab = 'Length')
```

## Third Dataset

Source: <https://datadryad.org/stash/dataset/doi:10.5061/dryad.50720>

This third dataset looks to determine if location has the same effect on average mass of fish as it does on the mass of land animals.

A question could be what is the effect of latitude.

```{r}
 fish <- read.csv("fish.csv")
summary(fish)
```
```{r, echo=FALSE}
boxplot(fish$Weight.maxima..g.~fish$Species,
        main = 'Mass of Fish Species',
        xlab = 'Species',
        ylab = 'Mass')
```
