---
title: "DMorrowWk4HW"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```
Destinee Morrow
Here is my Week 4 Homework

I found it quite difficult to find url links to the csv files so i could download straight from internet instead of saving to local drive.
Hoping we can go over reproducability when it comes to this. 

Could not find raw data or .csv URL link, so:
Download data into CSV from website
https://archive.ics.uci.edu/ml/machine-learning-databases/horse-colic/horse-colic.data

I found this data set interesting because my Bachelors is in Animal Science and I didnt work too much with horses, so I'd like to learn some more.
Some difficulties I think i may encounter is that some columns have alot of NA
I would like to determine whether or not PCV can determine the outcome of colic

```{r Horse Colic}
library(dplyr)
horsecolic <- read.csv("Documents/Programming/horsecolic.csv")
View(horsecolic)
```

This dataset had both quantatative and qualitative data.
Having a total of 28 different coloumns including their unique ID.
Here is a sample bar chart.

```{r horse, echo=TRUE}
library(ggplot2)
ggplot(horsecolic, aes(temp_of_extremities,rectal_temp), stat='mean') +geom_col()
```

This data is regarding white wine. I wanted to look at something completely different from the fist dataset i choose. 
This data has many more entries than that of the last and may make it harder to look at. However, there are less columns to look at.
This data set also does not have a unique idenifier for each wine.
I would like to determine if there is a correlation between alcohol content and sugars present.

```{r wine}
download.file('https://archive.ics.uci.edu/ml/machine-learning-databases/wine-quality/winequality-white.csv',"whitewine.csv")
dfwine <- read.csv('whitewine.csv', sep = ';', header = TRUE)
View(dfwine)
```

This dataset looks at various, 12 total, quantatative data.
Here is a sample scatterplot. 

```{r wine plot}
ggplot(dfwine, aes(citric.acid, chlorides)) +geom_point() + ggtile("Chlorides as a function Citric Acid in White wine") + xlab("Citric Acid") +ylab("Chlorides")
```

My third and final dataset i decided to look at is Minerals. I found this dataset interesting because ive always had an interest in gemstones and collecting them. 
This data may be difficult to work with because of the input of long strings into values. If i wanted to look for a specific element i would have to seach within those strings. 
I want to see if there is a type of element that shows up the most throughout all the elements.

```{r}
minerals <- read.csv("Documents/Programming/MineralsData.csv")
View(minerals)
```

This dataset has both quantative and qualatatives attritibutes. 
Here is a sample histogram. Obviously this needs cleaned up because you cant read the location names.

```{r}
ggplot(minerals, aes(Country.of.Type.Locality), stat = "bin") +geom_bar() +xlab("Location of Mineral") +ylab("Number of Minerals") + ggtitle("Numbers of Minerals found in Specific Location")

```
